Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-02-08 analysis/paper_draft3.Rmd
update Batson app screenshots, references, Discussion

Bias detection also increases with the number of historical trials and the number of strikes per trial. Higher values of $\alpha$ lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \@ref(fig:figcmptbly)). The choice of $\alpha$ matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \@ref(fig:figcmptbly)). Overall, given high compatibility, increasing $\alpha$ will increase bias detection. But our simulations also show that, given high incompatibility, bias detection can be erroneously high when current bias ($b_{curr}$) is low; and bias detection can be erroneously low where current bias is high but bias in historical trials ($b_{hist}$) is low.

@@

Suppose the same scenario discussed in section \@ref(prototype), except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors. If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)).

```{r figapp3, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Gender). Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney. Equal weight ($a = 1$) applied to historical and current data.", fig.align='center'}

load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot3.png")

@@

Here, most of the prosecutor's density curve of both prior and posterior are to the left of zero, indicating the possible bias against male jurors. However, because the 80\% credible interval includes zero, inferring such bias from the strike data alone is unjustified.

```{r figapp4, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Gender). Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney. Minimal weight ($a = 0.1$) applied to historical data.", fig.align='center'}

@@

Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure \@ref(fig:figapp4)). Both the prior and posterior density become flatter and the credible interval plainly includes zero.

>>>>

Bias detection also increases with the number of historical trials and the number of strikes per trial. Higher values of $\alpha$ lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \@ref(fig:figcmptbly)). The choice of $\alpha$ matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \@ref(fig:figcmptbly)). Overall, given high compatibility, increasing $\alpha$ will increase bias detection. But our simulations also show that, given high incompatibility, bias detection can be (erroneously) high when current bias ($b_{curr}$) is low; and bias detection can be (erroneously) low where current bias is high but bias in historical trials ($b_{hist}$) is low.

@@

Suppose the same scenario discussed in section \@ref(prototype), except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors. If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)(left)).

```{r figapp3, out.width='50%', fig.cap = "Screenshots of R-Shiny application show density plots for gender bias of prosecutor and defense attorney based on historical and current strike data. Vertical dotted lines depict 80 pct. credible interval.", fig.show = 'hold', fig.ncol = 2, fig.subchap=c("Equal weight","Minimal weight") }

load_path3 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot3.png")
load_path4 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot4.png")

#par(mfrow=c(2,1))

knitr::include_graphics(load_path3)
knitr::include_graphics(load_path4)

@@

Here, most of the prosecutor's density curve of both prior and posterior are to the left of zero, indicating the possible bias against male jurors. However, because the 80\% credible interval includes zero, inferring such bias from the strike data alone is unjustified. Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure \@ref(fig:figapp3)(right)). Both the prior and posterior density become flatter and the credible interval plainly includes zero.

```{r figapp4, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Gender). Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney. Minimal weight ($a = 0.1$) applied to historical data.", fig.align='center', eval=FALSE}

@@
